Title Text

### 2011-06-10 talk/iwtc11/paper.tex
description of heap caching

\subsection{Heap caching}

To implement heap caching variables has to be passed from the first
iteration to the second by XXX
\begin{equation}
  \hat J = \left(J_1, J_2, \cdots, J_{|J|}, H_1, H_2, \cdots, H_{|H}\right)
\end{equation}
\begin{equation}
  \hat K = \left(K_1, K_2, \cdots, K_{|J|}, m(H_1), m(H_2), \cdots, m(H_{|H})\right)
  .
\end{equation}
In the optimized trace $I$ is replaced by $\hat I$ and $K$ by $\hat K$.

>>>>

The guard on $p_5$ on line 17 of Figure~\ref{fig:unopt-trace} can be
removed since $p_5$ is allocated on line 10 with a known class. The
guard on $p_0$ on line 20 can be removed since it is identical to the
guard on line 6.

\subsection{Heap caching}
The objective of heap caching is to remove \lstinline{get} and
\lstinline{set} operations whose results can be deduced from previous
\lstinline{get} and \lstinline{set} operations. Exact details of the
process are outside the scope of this paper We will here assume that
it works perfectly and only consider the interactions with loop
peeling.

The issue at hand is to keep the second iteration a proper
trace. Consider the \lstinline{get} operation on line 19 of
Figure~\ref{fig:unopt-trace}. The result of this operation can be
deduced to be $i_4$ from the \lstinline{set} operation on line
12. Also, the result of the \lstinline{get} operation on line 22 can
be deduced to be $i_3$ from the \lstinline{get} operation on line
8. The optimization will thus remove line 19 and 22 from the trace and
replace $i_6$ with $i_4$ and $i_7$ with $i_3$.

After that, the second
iteration will no longer be proper as it operates on $i_3$ and $i_4$
which are not part of it. The solution is to extend the input
arguments, $J$, with those two variables. This will also extend the
jump arguments of the first iteration, which is also $J$. 
Implicitly that also extends the jump arguments of the second iteration, $K$,
since they are the inlined versions of $J$. That is the, $I$ has to
be replaced by $\hat I$ which is formed as a concatenation of $I$ and
$\left(i_3, i_4\right)$. At the same time $K$ has to be replaced by
$\hat K$ which is formed as a concatenation of $K$ and 
$\left(m\left(i_3\right), m\left(i_4\right)\right) = \left(i_7, i_8\right)$. 
The variable $i_7$ will then be replaced by $i_3$ by the heap caching
algorithm as it has removed the variable $i_7$. XXX: Maybe we should
replace $i_7=$get(...) with $i_7=i_3$ instead of removing it?

In general what is needed is for the heap optimizer is to keep track of
which variables from the first iterations it reuses in the second
iteration. It has to construct a vector of such variables $H$ which
can be used to update the input and jump arguments,
\begin{equation}
  \hat J = \left(J_1, J_2, \cdots, J_{|J|}, H_1, H_2, \cdots, H_{|H}\right)
\end{equation}
\begin{equation}
  \hat K = \left(K_1, K_2, \cdots, K_{|J|}, m(H_1), m(H_2), \cdots, m(H_{|H})\right)
  .
\end{equation}
In the optimized trace $I$ is replaced by $\hat I$ and $K$ by $\hat
K$. The trace from Figure~\ref{fig:unopt-trace} will be optimized into
